---
title: "Project A1 : Fungal invasion of the Apple fruit"
author: "Auriane Cozic, Thibault de La Taille, Eloi Littner, Audrey Ménaësse"
date: "May 1st, 2020"
output: 
  pdf_document:
    df_print: paged
geometry: margin=2.5cm
fontsize: 12pt
df_print: paged

header-includes:
- \usepackage{float}
- \usepackage{fancyhdr}
- \usepackage{wrapfig, subfig}
- \pagestyle{fancy}
- \fancyhead[C]{Cozic, De La Taille, Littner, Ménaësse}
- \fancyhead[L]{}
- \fancyhead[R]{}

---



# Introduction

Apples are among the best-preserved fruits. However, after a while fungi strains are inevitably starting to grow. In F.G. Gregory and A.S.Horne study, the conditions of infection, the progression of the invasion and the final stage of invasion were studied for two varieties of apple : the Cox’s Orange Pippin from Burwell, Cambridgeshire and Bramley’s seeding from 6 different localities. The data set under study describes the fungal invasion of 35 apples. It contains 7 different variables, regarding:

* the characteristics of the apple (variety, weight, radius)
* the strain of the fungi, separated in 7 types (A, B11, B111, C1, C21, C3 and D)
* the measurements of the infection (days after infection, fungal radial advance, rate of fungal advance).

The aim of our study is to compare the infecting power depending on the conditions of infection and the resistance to invasion of different varieties of apples and their characteristics.





```{r, include=FALSE}
#Load libraries and packages
require(MASS)
library(ggplot2)
library(plotrix)
```

```{r, include=FALSE}
data <-read.table("http://users.stat.ufl.edu/~winner/data/apple1.dat")
colnames(data) <- c("variety", "strain", "days", "weight", "radius", "fungal radial advance", "rate of advance")
```

# Exploratory data analysis
## Data description

To begin with, we want to get familiar with the data. The variable *variety* contains several metadata information. Therefore, in order to have explicit variables in our model, we decided to split it into two new variables : the year when the experiment was conducted and the storage temperature of the apples. 

The study is looking at 2 different varieties of apples and 7 fungi strains. It comes from 5 different experiments started in 1924 or 1925. 4 of them were performed at the same temperature (12°), but one was at (3°). The duration between the infection and the measurements is different for each experiment.



```{r, include=FALSE}
data$variety[data$variety <= 2]<-"Bramley"
data$variety[data$variety <= 5]<-"Cox"
data$year = 1924
data$year[data$days == 70] <- 1925
data$year[data$days == 86] <- 1925
data$temperature<-12
data$temperature[data$days == 138]<-3
data$variety <- as.factor(data$variety)
data$strain <- as.factor(data$strain)
data$year <- as.factor(data$year) 
data$temperature<- as.factor(data$temperature)
```




## Outliers

Exploring the data, 3 outliers can be detected.
The Cox apples *18* and *21* have a fungal radial advance superior to the apple radius. As it does not have any physical sense or meaningful explanation, they were removed.
The infection did not develop in one apple *17*. The fungal radial advance is very lower to any other apple. As it the only case, it reveals an experimental error and was also removed.

 
```{r, include=FALSE}
data.rm<- data[-c(17,18,21),]
```



#Correlations -> remove days and radius ; consider only rate of advance (not radius) 
#TODO : replace data by data.rm and rerun 



Explore correlations and colinearity: 
```{r}
pairs(data[c("days", "weight", "radius", "fungal radial advance", "rate of advance")])
#cov(data[c("days", "weight", "radius", "fungal radial advance", "rate of advance")])
cor(data[c("weight", "radius", "fungal radial advance", "rate of advance")])
#plot(x = data$days, y=data$year)
require('ggplot2')
layout(matrix(1:3, ncol=3))
qplot(data$days,data$year)
qplot(data$days,data$temperature)
qplot(data$days,data$variety)
# ""
# "corrected correlations"
# ""
# 
# p.cor<-function(x){
#   inv<-solve(var(x))
#   sdi<-diag(1/sqrt(diag(inv)))
#   p.cor.mat<--(sdi%*%inv%*%sdi)
#   diag(p.cor.mat)<-1
#   rownames(p.cor.mat)<-colnames(p.cor.mat)<-colnames(x)
#   return(p.cor.mat)}
# #p.cor(data[c("strain","days", "weight", "radius","temperature", "fungal radial advance", "rate of advance", "year")])
```
#TODO : virer le matrix plot, et voir si on a la place de mettre les trois graphs de la fin, sinon les t?j et mettre une phrase. Move to supplementary 
On voit bien une forte corrélation entre "weight" et "radius" sur le 1er graphe, ce qui est confirmé par les coefficients de corrélation; de même entre "rate of advance" et "fungal radial advance". 
Cela est bien cohérent en fonction des méthodes de calcul liant les 2 respectivement (avec approximation pour weight). 

De plus, en observant les relations entre "days" et les autres facteurs ("year", "temperature" et "variety", d'où sont extraites ces informations), il apparaît que "days" contient toutes ces facteurs, mais de manière non linéaire; et est donc à exclure car risquant de fausser l'analyse sans gain d'information.


#Exploratory data analysis 

```{r, fig.width=6, fig.height=4}
plot.design(x=data.rm$`rate of advance`~.,data=data.rm[,c(1,2,7,8,9)], ylab= 'Rate of fungal advance (cm/day)', main = "Univariate effects of factors", cex=0.7, cex.lab=0.8 ,cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
#cex changes font size within the graph to prevent overlaps
```
The strain, temperature and year seem to have a strong influcence on the rate of fungal advance, while the variety has low impact. 

###if enough space, graphs
This can be confirmed by looking at the data points

```{r}
par(mfrow=c(2,2))
bp1 <- boxplot(data.rm$`rate of advance`~data.rm$variety, ylab='Fungal rate of advance (cm/day)', xlab = "Variety")
text(bp1$group, bp1$out, rownames(data.rm[data.rm$`rate of advance` %in% bp1$out,]),pos = 4, cex=0.7) #label the outliers
bp2 <- boxplot(data.rm$`rate of advance`~data.rm$strain, ylab='Fungal rate of advance (cm/day)', xlab = "Fungal strain")
text(bp2$group, bp2$out, rownames(data.rm[data.rm$`rate of advance` %in% bp2$out,]),pos = 4, cex=0.7)
bp3 <- boxplot(data.rm$`rate of advance`~data.rm$temperature, ylab='Fungal rate of advance (cm/day)', xlab = "Temperature (°)")
text(bp3$group, bp3$out, rownames(data.rm[data.rm$`rate of advance` %in% bp3$out,]),pos = 4, cex=0.7)
bp4 <- boxplot(data.rm$`rate of advance`~data.rm$year, ylab='Fungal rate of advance (cm/day)', xlab = "Year")
text(bp4$group, bp4$out, rownames(data.rm[data.rm$`rate of advance` %in% bp4$out,]),pos = 4, cex=0.7)
```

The fungi strains of type V (2 and 3), and a cold temperature, seem to have a reduced infecting power.
It can also be noticed that Cox variety, fungi strain 1, 12° temperature and year 1924 experiments display higher variability, which is confirmed by their standard deviations. This is due to the fact that each of these categories gather only few data (less than 5). Any generalisation of the results is thus complicated.


##Do not show
```{r, include=FALSE}
tapply(data.rm$`rate of advance`, list(data.rm$strain), sd)
tapply(data.rm$`rate of advance`, list(data.rm$variety), sd)
tapply(data.rm$`rate of advance`, list(data.rm$year), sd)
tapply(data.rm$`rate of advance`, list(data.rm$temperature), sd)
```


3 outliers apples stand out having particularly high Fungal rate of advance (15, 19 and 20), and one has a slightly lower one (10).

###graph only if enough space
```{r}
plot(data.rm$weight, data.rm$`rate of advance`, xlab = "Weight (g)", ylab = "Rate of fungal advance (cm/day)", main = "Effect of the apple weight on the infection")
```
Finally, we can not see any pattern on the rate of fungal advance depending on the weight, neither abnormal variability. 


####Model selection

Definition of the different linear models




##Stepwise selection of variables
```{r}
fit.stepAIC <- stepAIC(lm(`rate of advance` ~ variety+strain+weight+temperature+year, data=data))
```


#TODO : not show above, only that year was removed (and show further that coeffs are roughly the same)


```{r}
fit.stepAIC
```


#Interpretation

```{r,include=F}
for (k in c(4)) {
  plot(fit.stepAIC, which=k)
}
```


#If enough space, add some elements to compare the models
```{r, include=F}
fit.rm.stepAIC <- stepAIC(lm(`rate of advance` ~ variety+strain+weight+temperature, data=data.rm))
fit.rm.stepAICfull <- stepAIC(lm(`rate of advance` ~ variety+strain+weight+temperature+year, data=data.rm))
summary(fit.rm.stepAIC)
```


#TODO : interpret the model, explain the outliers 
#TODO : explain here why we don't fit the full model (remove year) : coherent with full data (with outliers)

```{r}
anova(fit.rm.stepAIC,fit.rm.stepAICfull)
```




# Assessment of the model

In order to assess the model found previously, we first check the QQ-plot of the fitted data. 
```{r echo = F, warning = F, message = F, fig.width=7, fig.height = 7, fig.cap = "Diagnostic plots", fig.subcap=c("\\label{fig:QQ}QQ-plot", "\\label{fig:residuals}Residuals"), fig.pos= "htbp", out.height="8cm", out.width=".8\\linewidth", fig.align="center"}
 plot(fit.rm.stepAIC, which=2, xlab= 'Theoretical quantiles of the linear model', sub="") #QQ
 plot(fit.rm.stepAIC, which=1, sub="")#residuals
```

The tail of residuals is composed of a very few data points which are significantly far from the model. The data set is then well represented by the linear model.

The study of the residuals is also important to assess the model. Indeed, if the residuals are not carrying any information relevant to the model, their mean should be zero which is almost the case in the following plot.






##Check for time dependencies (Thibault)

test time dependence of rate of advance 
```{r}
plot(x=data.rm$days, y= residuals(fit.rm.stepAIC))
title("residuals= f(days)")
```

Homoscedasticity among all groups


#Discussion / Results (short)


###Conclusion
#TODO : finish it
The most impactant variable is the day of the infection : the infection is strongest (progresses more rapidly) when the apple is infected early.
(Pas tres sure de comment interpreter ça ?)

The strains B (strains 2 and 3) are very less virulent.

Finally, low temperature leads to more resistance of the apple.

#Question to the teacher
Is our method ok or only ANOVA ?
Do you validate the part when we remove year ? 


#R?partition

Corr?lations + Conclusion : Thibault 
EDA sauf corr?lations : Audrey 
Fit du model : Eloi
Residuals + discussion : Auriane

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
