---
title: "Project A1 : Fungal invasion of the Apple fruit"
author: "Auriane Cozic, Thibault de La Taille, Eloi Littner, Audrey Menaesse"
date: "March 20, 2020"
output: 
  pdf_document:
    df_print: paged
geometry: margin=2.5cm
fontsize: 12pt
df_print: paged
---


# Data description


Description : This data set is used to describe the fungal invasion of 5 varieties of apple by 7 different fusarium strains.  

Variables/names :

Questions : - Is radius the radius of the apple ? If yes, how do you interpret the fact that the Fungal Radial Advance is bigger than the size of the apple ?

#Import and loading data

```{r, echo=FALSE}
#Load libraries and packages
```


```{r}
data <-read.table("http://users.stat.ufl.edu/~winner/data/apple1.dat")
colnames(data) <- c("variety", "strain", "days", "weight", "radius", "fungal radial advance", "rate of advance")
data
```


# Exploratory data analysis 

```{r}
summary(data)
```
Different plots to illustrate the data :
```{r}
plot(data$strain,data$`rate of advance`)
plot(data$variety,data$`rate of advance`)
plot(data$days,data$`rate of advance`)
plot(data$strain, data$`fungal radial advance`)
plot(data$variety, data$`fungal radial advance`)
plot(data$days, data$`fungal radial advance`)
```

Here, we want to observe the rate of advance of the fungal invasion depending on the fusarium strain and the variety of apple. 
```{r}
tapply(data$`rate of advance`, list(data$strain, data$variety), mean)
```

Now we want to compare the rate of advance with the number of days
```{r}
tapply(data$`rate of advance`, list(data$strain, data$days), mean)
```

Definition of the different linear models
```{r}
fit.s = lm(`rate of advance` ~ strain, data=data)
fit.sv = lm(`rate of advance` ~ strain+variety, data=data)
fit.svd = lm (`rate of advance` ~ strain+variety+days, data=data)
fit.svdw = lm (`rate of advance` ~ strain+variety+days+weight, data=data)
fit.svdwr = lm (`rate of advance` ~ strain+variety+days+weight+radius, data=data)

```

Comparing the different models with anova method :
```{r}
anova(fit.s, fit.sv)
```
